High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales

### 2021-03-04 inst/manuscript/manuscript.Rmd
Edited through discussion, abstract still missing

@@

documentclass: article
classoption: a4paper
capsize: normalsize
fontsize: 11pt
geometry: margin=1in
linenumbers: yes
spacing: doublespacing
footerdate: yes
abstract: The abstract can go either here or below
keywords: rmarkdown, reproducible science
bibliography: 
  - references.bib
csl: ecology.csl

@@

The second requirement is addressed by a computer vision/machine learning classifer that is effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016], re-engineered to run on high performance, cloud-based computing clusters, and following recommended practices for controlling for and measuring the errors that occur when training machine learning models for remote sensing applications [@ElmesAccountingtrainingdata2020]. The classifier is tightly coupled to a front-end platform for collecting label data, which includes rigorous accuracy assessment protocols and a novel approach for merging multiple maps into consensus labels, thereby minimizing image interpretation error [@Estesplatformcrowdsourcingcreation2016; @ElmesAccountingtrainingdata2020]. The training and machine learning components are combined within an "active learning" [@cohnImprovingGeneralizationActive1994] framework, wherein the machine learning process assesses classification uncertainty in unlabelled areas after a training step, and selects sites from areas of highest uncertainty for additional labelling [@DebatsIntegratingactivelearning2017]. Our framework automates this interactive approach to label selection, which has been shown to be important for improving accuracy in previous studies [e.g. @XiongNominal30mCropland2017]. Furthermore, it is more efficient than purely randomized or conventionally stratified approaches to label collection, meaning that fewer labels are needed to generate a given level of accuracy [@cohnImprovingGeneralizationActive1994; @DebatsIntegratingactivelearning2017]. Finally, an unsupervised segmentation step is applied to the imagery and merged with the pixel-wise classifications from the machine learning process, resulting in a final vector-based field boundary map that provides important information on field geometry.

@@

```{r, message = FALSE, warning = FALSE, echo=FALSE}
prdelta <- iteration_metrics %>% 
  filter(Metric %in% c("FPR", "TPR")) %>% 
  group_by(Iteration, Metric) %>% 
  summarize(Score = mean(Score)) %>% 
  filter(Iteration %in% c(0, 3)) %>% 
  tidyr::pivot_wider(names_from = Iteration, values_from = Score) %>% 
  mutate(rate = round((`0` - `3`) / `0` * 100, 1))
```
The active learning approach helped to improve overall model performance relative to randomized training site selection, confirming previous findings reported by @DebatsIntegratingactivelearning2017. Although the performance gains relative to randomized model training that we observed were substantially less than in the previous study, which reported `r round((0.45-0.35) / 0.35 * 100)`% higher model performance after one iteration, and `r round((0.69-0.64) / 0.64 * 100)`% higher after 45 iterations [@DebatsIntegratingactivelearning2017], those comparisons were made from a low initial base, beginning with a single training site with one more site added each iteration. In contrast, our large initial random sample (500) meant that our models were substantially trained before they were exposed to actively selected labels, thereby diluting their impact on performance. Nevertheless, the higher average performance of the active approach across three performance metrics demonstrated its effectiveness. Most notable were the larger gains in the F1 score (Figure S7), a balanced performance metric. The gains in accuracy and AUC were smaller. For accuracy, the reduced advantage was primarily due to active learning being outperformed by randomized training after the first iteration in AOI 15, which proved one of the hardest AOIs to both map and label. Active learning likely resulted in the selection of sites that were harder to label than randomly selected ones, leading to more label error, and thus initially lower model accuracy. However, this deficit was overcome by the 5th iteration. The plateau in AUC gains at 0.5% better than randomized training reflects the findings that active learning reduced both the false and true positive rates, the two inputs to AUC. Although the false positive rate declined by `r prdelta$rate[1]`% between Iterations 0 and 3, the `r prdelta$rate[2]`% reduction in the true positive rate offset some of these gains, and the effect of this tradeoff may have been more pronounced in the threshold independent AUC.

The detail, temporal precision, and large extent of these maps was enabled by the system's ability to process PlanetScope data, which is currently the only source of sub-5 meter imagery with daily coverage [@McCabefutureEarthobservation2017]. Daily revisits are necessary for creating seasonal composites within a single agricultural year over cloudy areas. The compositing technique we developed to work with PlanetScope was able to provide complete image coverage over the entire country for both seasons. Although Sentinel-2 is free, has better radiometric quality, and has sufficiently high resolution (10 m) to accurately classify small-scale agricultural systems [e.g. @Defournyrealtimeagriculturemonitoring2019; @kernerRapidResponseCrop2020], its 5-day interval is too infrequent to generate sufficiently cloud-free composites during the growing season over southern Ghana. Sentinel-1 is not affected by the same problem, but labelling fields in more coarsely resolved radar images is difficult.

## Lingering questions
Several potential issues not answered by our assessment merit further exploration. One of these is the degree of correspondence between image- and ground-collected labels. However, such comparisons will likely reveal unresolvable discrepancies in most circumstances. The highly dynamic nature of the systems in question means that even narrow temporal discrepancies between ground and image-collected labels may show substantial differences. These discrepancies could be further exacerbated by differences in an understanding of what constitutes a field, which might vary on the ground depending on the farmer or the data collecter's collectors judgement. On top of this, field-collected boundaries are typically highly geographically constrained and limited in size, meaning that any comparison might suggest that the relationship is more noisy than it actually is. These factors suggest that ground versus image label differences would not necessarily indicate how far image-based labellers were from the "truth". Nevertheless, a comparison against ground data can help to assess how accurately image-collected labels capture the typical size of fields, and thus merits further investigation for this reason.

The temporal discrepancies mentioned above [and discussed in @ElmesAccountingtrainingdata2020] are another reason why we chose not to  label on basemap imagery (in addition to restrictive usage terms), which is typically several years old [@LesivCharacterizingspatialtemporal2018]. However, we did not assess whether the higher label accuracy one might achieve by digitizing on <1-2 m basemap would offset model errors due temporal mismatches between labels and imagery.

Another potential issue is the degree to which our assessment of the impact of label error on model performance (Figure \ref{fig:trainingimpact}) was influenced by the validation labels we used, which were generated using the consensus method. This could have confounded the assessment, particularly when comparing models trained with the most accurate individual label and those trained with consensus labels. However, the visual assessment of probability images generated from each labelling strategy show that these two strategies produce nearly identical maps, and that the model trained with low quality labels generated a markedly less certain map (Figure S6).

@@

## Conclusion
This work demonstrates the proof of concept for developing high resolution, annual maps of small-scale croplands at national to regional scales, using a framework that can be readily updated to improve map accuracy as technologies improve. Such maps will be important for developing an understanding of the rapid agricultural change that is currently unfolding throughout much of the continent.  
# Acknowledgements
The primary support for this work was provided by Omidyar Network's Property Rights Initiative, now PlaceFund. Additional support was provided by NASA (80NSSC18K0158), and the National Science Foundation (SES-1801251; SES-1832393). Computing support was provided by the AWS Cloud Credits for Research program and the Amazon Sustainability Data Initiative. Azavea provided significant support in developing the machine learning pipeline. We thank Meridia for assistance in initial data about local cropping systems and the characteristics of fields, and Radiant Earth Foundation for advice and guidance regarding machine learning best practices.

>>>>

- name: Angeline Hokar
  affilnum: '3'

- name: Charles Juma 
  affilnum: '3'

- name: Jackson Makuru
  affilnum: '3'

- name: Fredrick Matasa
  affilnum: '3'

- name: Ben Mbatia
  affilnum: '3'

- name: Judith Mbaye
  affilnum: '3'

- name: Isaac Muasa
  affilnum: '3'

- name: Julius Mucha
  affilnum: '3'

- name: Adelide Mugami
  affilnum: '3'

- name: Purent Oduor
  affilnum: '3'

- name: Jeff Ochieng
  affilnum: '3'

- name: Francis Wambua
  affilnum: '3'

- name: Thuo Wanjiku
  affilnum: '3'

- name: Joe Wanyoike 
  affilnum: '3'

@@

documentclass: article
classoption: a4paper
capsize: normalsize
fontsize: 11pt
geometry: margin=1in
linenumbers: yes
spacing: doublespacing
footerdate: yes
abstract:

keywords: rmarkdown, reproducible science
bibliography: 
  - references.bib
csl: ecology.csl

@@

The second requirement is addressed by a computer vision/machine learning classifer that is effective for classifying smallholder croplands [@Debatsgeneralizedcomputervision2016], re-engineered to run on high performance, cloud-based computing clusters, and following recommended practices for controlling for and measuring the errors that occur when training machine learning models for remote sensing applications [@ElmesAccountingtrainingdata2020]. The classifier is tightly coupled to a front-end platform for collecting label data, which includes rigorous accuracy assessment protocols and a novel approach for merging multiple maps into consensus labels, thereby minimizing image interpretation error [@Estesplatformcrowdsourcingcreation2016; @ElmesAccountingtrainingdata2020]. The training and machine learning components are combined within an active learning framework, wherein the machine learning process assesses classification uncertainty in unlabelled areas after a training step, and selects sites from areas of highest uncertainty for additional labelling [@cohnImprovingGeneralizationActive1994; @tuiaSurveyActiveLearning2011]. Our framework automates this interactive approach to label selection, which is effective in boosting the performance of classification models while reducing the overall number of training samples required to achieve a given level of performance [e.g. @hamrouniLocalGlobalTransfer2021; @DebatsIntegratingactivelearning2017]. Finally, an unsupervised segmentation step is applied to the imagery and merged with the pixel-wise classifications from the machine learning process, resulting in a vectorized field boundary map that provides important information on field geometry.

@@

```{r, message = FALSE, warning = FALSE, echo=FALSE}
prdelta <- iteration_metrics %>% 
  filter(Metric %in% c("FPR", "TPR")) %>% 
  group_by(Iteration, Metric) %>% 
  summarize(Score = mean(Score)) %>% 
  filter(Iteration %in% c(0, 3)) %>% 
  tidyr::pivot_wider(names_from = Iteration, values_from = Score) %>% 
  mutate(rate = round((`0` - `3`) / `0` * 100, 1))
```
The active learning approach helped to improve overall model performance relative to randomized training site selection, in line with findings from two recent efforts [@DebatsIntegratingactivelearning2017; @hamrouniLocalGlobalTransfer2021]. Although the performance gains relative to randomized model training that we observed were smaller (e.g. @DebatsIntegratingactivelearning2017 `r round((0.45-0.35) / 0.35 * 100)`% higher model performance after one iteration, and `r round((0.69-0.64) / 0.64 * 100)`% higher on the final iterations), those comparisons were made from lower initial bases, with initial training samples that were less than 1/10 the size, in terms of pixels, of our initial training sample. Our large initial randomly selected sample (500 grid cells) meant that our models were substantially trained before they were exposed to actively selected labels, thereby diluting their impact on performance. Nevertheless, the higher average performance of the active approach across three performance metrics demonstrated its effectiveness. Most notable were the larger improvements seen in the F1 score (Figure S7), a balanced performance metric. Gains in accuracy and AUC were smaller. For accuracy, the reduced advantage was primarily due to active learning being outperformed by randomized training after the first iteration in AOI 15, which proved one of the hardest AOIs to both map and label. Active learning likely resulted in the selection of sites that were harder to label than randomly selected ones, leading to more label error, and thus initially lower model accuracy. However, this deficit was overcome by the 5th iteration. The plateau in AUC gains at 0.5% better than randomized training reflects the findings that active learning reduced both the false and true positive rates, the two inputs to AUC. Although the decline of the false positive rate (`r prdelta$rate[1]`% between Iterations 0 and 3) was nearly three times larger than that of the true positive rate (`r prdelta$rate[2]`%), AUC should be quite sensitive to the reduction in the latter, as it assesses how the tradeoff between the two rates varies across a full range of possible classification thresholds [@Pontiustotaloperatingcharacteristic2014].

The detail, temporal precision, and large extent of these maps was enabled by the system's ability to process PlanetScope data, which is currently the only source of sub-5 meter imagery with daily coverage [@McCabefutureEarthobservation2017]. Daily revisits are necessary for creating seasonal composites within a single  year over cloudy areas. The compositing technique we developed allowed us to develop a complete image catalog for the country representing the two seasons for 2018 agricultural year. Although Sentinel-2 is free, has better radiometric quality, and has sufficiently high resolution (10 m) to accurately classify small-scale agricultural systems [e.g. @Defournyrealtimeagriculturemonitoring2019; @kernerRapidResponseCrop2020], its 5-day interval is too infrequent to generate sufficiently cloud-free composites during the growing season over southern Ghana. Sentinel-1 is not affected by the same problem, but labelling fields in more coarsely resolved radar images is challenging.

## Lingering questions
Several potential issues not answered by our assessment merit further exploration. One of these is the degree of correspondence between image- and ground-collected labels. However, such comparisons will likely reveal unresolvable differences between the two perspectives. The highly dynamic nature of many agricultural systems means that relatively narrow differences between the dates of ground- and image-based digitizing campaigns can lead to substantial disagreement between the resulting field boundaries, simply because the fields themselves may have shifted during the interval [@ElmesAccountingtrainingdata2020]. These discrepancies could be further exacerbated by differences in the definition of what constitutes a field, which might vary on the ground depending on who is being asked, or who is doing the collecting. On top of this, ground-based collection efforts are usually highly geographically constrained and limited in size, meaning that any comparison might suggest that the relationship is more noisy than it actually is, or may be unrepresentative. These factors suggest that ground versus image label differences would not necessarily indicate how far image-based labellers were from the "truth". Nevertheless, a comparison against ground data would help to assess how accurately image-collected labels capture the typical size of fields, and thus merits further investigation for this reason.

The temporal discrepancies mentioned above [and discussed in @ElmesAccountingtrainingdata2020] are another reason why we chose not to label on basemap imagery (in addition to restrictive usage terms), which is typically several years old [@LesivCharacterizingspatialtemporal2018]. However, we did not assess whether the higher label accuracy one might achieve by digitizing on a <1-2 m resolution basemap would offset model errors caused by temporal mismatches.

Another potential issue is the degree to which our assessment of the impact of label error on model performance (Figure \ref{fig:trainingimpact}) was influenced by the validation labels we used, which were generated using the consensus method. This could have confounded the assessment, particularly when comparing models trained with the most accurate individual label and those trained with consensus labels. However, the visual assessment of their resulting probability maps confirm the differences in scores: consensus and most accurate individual labels produce nearly identical maps with relatively high certainty, while low quality labels led to a markedly less certain map (Figure S8).

@@

## Conclusion
This work demonstrates the proof of concept for developing high resolution, annual maps of small-scale croplands at national to regional scales, using a framework that can be readily updated to improve map accuracy as technologies improve. Such maps will be important for developing an understanding of the rapid agricultural change that is currently unfolding throughout much of the continent.

# Acknowledgements
The primary support for this work was provided by Omidyar Network's Property Rights Initiative, now PlaceFund. Additional support was provided by NASA (80NSSC18K0158), and the National Science Foundation (SES-1801251; SES-1832393). Computing support was provided by the AWS Cloud Credits for Research program and the Amazon Sustainability Data Initiative. Azavea provided significant contributions in engineering the machine learning pipeline. We thank Meridia for providing initial information about local cropping systems and the characteristics of fields, and Radiant Earth Foundation for advice and guidance regarding machine learning best practices.
